---
title: "credit_card_clustering"
author: "Tai Vo"
date: "2024-06-07"
output: html_document
---

#Import packages
```{r}
library(Hmisc) #Correlation test
library(corrplot) 
library(scatterplot3d)
library(factoextra)
library(NbClust)
library(tidyverse)
library(cluster)
library(reshape2)
```

##Load dataset
```{r}
credits1 <- read_csv("CC GENERAL.csv")
```

##Check null values
```{r}
colSums(is.na(credits1))
```
##Drop null values
```{r}
credits <- drop_na(credits1)
```

##Use 3 columns: BALANCE, PURCHASE, and CREDIT_LIMIT to conduct analysis
```{r}
data <- as.data.frame(credits[, c("BALANCE", "PURCHASES", "CREDIT_LIMIT")])
head(dat)
```

##Pearson correlation test
```{r}
crlt1 <- cor(dat, method = "pearson")
crlt1
```

##Correlation visualization
```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(crlt1, method = "color", col = col(200),  
         type = "upper", order = "hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col = "black", tl.srt = 45, #Text label color and rotation
)
```
##Hopkins' stat coefficient testing
Use Hopkins' statistic measure to check whether the dataset is clusterable(close to 1)
```{r}
cl = get_clust_tendency(dat, n=40, graph=T)
summary(cl)
```
## Draw dendrogram with 7 cluster 
```{r}
d <-dist(dat, method ="euclidean")
hc <- hclust(d, method = "ward.D2")
plot(hc)
rect.hclust(hc, k=7, border="red")
```


## Select k methods:
```{r}
# Elbows method
fviz_nbclust(dat, kmeans, method="wss") + geom_vline(xintercept = 7, linetype = 5)
```

## R Markdown
```{r}
# Average Silhouette method
fviz_nbclust(dat, kmeans, method="silhouette", k.max = 30, nboot = 10) + geom_line()
```

## R Markdown
```{r}
#Gap statistics method
set.seed(123)
fviz_nbclust(dat, kmeans, k.max = 10, nstart=25, nboot=50, method="gap_stat") + geom_line()
```

##Results of K-means clustering with 7 clusters
```{r}
km = kmeans(dat, centers = 7, nstart = 20)
km
```

##Evaluate the clustering results of the model
Use silhouette measure (how similar an objects is to the other objects in its own cluster )
Average Silhouette width range from -1(poor clustered) to 1(good clustered)
```{r}
hc_sil = silhouette(km$cluster, dist(dat), method = "euclidean", lable = FALSE)
fviz_silhouette(hc_sil, print.summary = FALSE) + 
  theme_minimal()
```

##Create factor for clusters
```{r}
dat$cluster <- as.factor(km$cluster)
```

##Create color gradient
```{r}
colors <- c("royalblue1", "darkcyan", "oldlace","khaki","green", "plum", "lightblue")
```

##Visualize clusters using scatterplot3d function
```{r}
scatterplot3d(dat[,1:3], 
              main="Clusters Visualization",
              xlab = "Balance",
              ylab = "Purchases",
              zlab = "Credit Limit",
              pch = 16, 
              color=colors[as.numeric(dat$cluster)],
              box=FALSE)
legend("right", 
       legend = levels(dat$cluster),
       col = colors,
       pch = 16)
```

#Boxplot by clusters 
Show the value distribution of each variable in each cluster
```{r}
dat = melt(dat, id.var = "cluster")
ggplot(dat, aes(x = variable, y = value)) +
  geom_boxplot(aes(fill = cluster), outlier.size = 1) +
  facet_wrap( ~ variable, scales = "free", ncol = 3, nrow = 2) +
  labs(x = NULL, y = NULL) +
  theme_minimal()
```


